\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-22 k-planar.tex
Pfffft
Lots of stuff
Working through......

These theorems have applications in diverse areas, which we explore in \cref{Applications}.

@@

Theorems~\ref{PlanarProduct}(b) and \ref{ShortcutProduct} and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$. Some further observations presented in \cref{sec-k-planar} lead to the improved result.  \cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs. Some examples are explored in \cref{first_example,mid_example,last_example}.

@@

@@

The purpose of this section is to prove our main technical result, \cref{ShortcutPartition}, and its specialization to $k$-planar graphs \cref{kPlanarProduct}.

\subsection{Product Structure for Shortcut Systems}

\cref{ShortcutPartition} shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, an $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.

For convenience, it will be helpful to assume that $\SS$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.  Since $G^\SS$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\SS}$ but eliminates special cases in some of our proofs.

@@

We are now ready to prove our main result.
% , which we restate here for convenience:

@@

\subsection{\boldmath Product Structure for $k$-Planar Graphs}

@@

@@

A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar. Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b) and \cref{StringShortcut,qn} imply:

@@

\section{Framed Graphs, $1$-Planar Graphs, and Map Graphs}

@@

For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane \note{DW}{Use the language of ``embedded graph''} having a biconnected plane spanning (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$. The graph $G_0$ is called the \defin{frame} of $G_0$. This definition (for simple graphs) was introduced by \citet{BDGGMR}. Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar multigraph (by adding a vertex inside each face $F$ adjacent to the vertices of $F$). Thus \cref{ShortcutProduct} is applicable. However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}

@@

This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs.
For simple 1-planar graphs it is essential that we allow frames with parallel edges.

@@

\note{PM}{I would also like to change $G'$ to $G_0$ and $N'$ to $N_0$.  One issue is holding me back, though: We recursive on cycles $F_1,\ldots,F_m$ that define subgraphs $N_1,\ldots,N_m$ of $G$ and subgraphs $N_1',\ldots,N_m'$ of $G'$.  Will these latter subgraphs becomes $N_{0,1},\ldots,N_{0,m}$?}

\note{DW}{We need to define ``closure''}

@@

$1$-planar graphs are of particular interest to the graph drawing community~\citep{kobourov.liotta.ea:annotated}. In this case, we obtain better constants and an additional property (planarity) of $H$.

@@

\begin{lem}\label{1_planar_is_4_framed}
A graph $G$ is $1$-planar if and only if it is contained in a $4$-framed multigraph.
\end{lem}

\note{DW}{Write  ``A multigraph $G$ is $1$-planar if and only if ...''}

@@

\begin{proof}
Let $G$ be a $1$-plane graph. We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.  This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.

@@

@@

%%%%%%%%%%%%%%%%%%%
\section{Allowing Crossings}
\label{sec-k-planar}

This section applies our main results for shortcut systems to prove a product structure theorem for graphs drawn with a bounded number of crossings per edge. Then we show how to improve the bounds in this case.

The definition of $k$-planar graphs naturally generalises for other surfaces. A graph $G$ drawn on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.  A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$. \cref{AddDummy} immediately generalises as follows:

Theorems~\ref{ShortcutProduct}  and \ref{GenusProduct}(b) imply a product structure theorem for $(g,k)$-planar graphs. The obtained bounds are improved by the following result, which is proved using exactly the same approach used in the proof of \cref{kPlanarProduct} (applying \cref{GenusProduct}(b) instead of \cref{PlanarProduct}(b)). We omit repeating the details.

Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{Examples}) was in terms of layered treewidth, which we now define.  A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$. The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.  The \defin{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$. \citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph. \citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$. It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of \citet{FP08}. In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).

@@

Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.

\referee{1}{4. Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to ``prove product structure theorems for several non-minor-closed classes of interest.'' Which of Corollaries 1--4 and Theorems 12--17 are most important? I would remove any mention of the applications from the examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure. The applications section is mostly a survey and I think should be focused on the most exciting new corollaries. Theorems 18 and 19 should be in the applications section.}

\note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
\note{PM}{Sure, though I think they proposing a chop rather than a restructuring.}

@@

\cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
% \note{OM}{Ok}
Note that \citet{DJMMUW20} previously proved the bound of
$O(g^{k+2})$ using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$
% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
% \note{PM}{Ok}
In the case of 1-planar graphs we can improve further. \citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$. Thus the graph $H$ in \cref{1-planar} has queue-number at most $5$. \cref{qn,1-planar} then imply:

@@

@@

\cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:

\begin{cor}\quad
\begin{compactitem}
\item	For every $1$-planar graph $G$, $\pi(G)\le 7\times 4^4=1792$.
%\item For every $k$-planar graph $G$,
%	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
\item	For every $(g,k)$-planar graph $G$,
	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
% \note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
% \note{PM}{Ok}
\end{compactitem}
\end{cor}

@@

@@

As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs. The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.

\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
\label{GenusProduct}
Every graph of Euler genus $g$ is contained in:
\begin{compactenum}[(a)]
\item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
\item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
\end{compactenum}
\end{thm}

>>>>

% A planar graph of treewidth at most $3$ is said to have \emph{simple treewidth} $3$.

For at least one application, the planarity of $H$ is critical in obtaining asymptotically tight bounds \cite{BDJM}. \cref{1_planar_product} is proven in the more general setting of $d$-framed multigraphs. This more general result implies a product structure theorem for $d$-map graphs in which $H$ is also planar.

% 
% 
% These theorems have applications in diverse areas, which we explore in \cref{Applications}.

@@

Theorem~\ref{PlanarProduct}(b), \cref{ShortcutProduct}, and \cref{AddDummy} imply \cref{kPlanarProduct} with $K_{6(k^3+3k)}$ instead of $K_{18k^2+48k+30}$. Some further observations presented in \cref{sec-k-planar} lead to the improved result.  \cref{ShortcutProduct} is applicable for many graph classes in addition to $k$-planar graphs. Some examples are explored in \cref{first_example,mid_example,last_example}.

@@

\subsection{Outline}

The remainder of this paper is organized as follows: 
\begin{compactitem}
  \item In \cref{Structure} we prove our main result for shortcut systems (\cref{ShortcutProduct}), and its optimisations for $k$-planar graphs and their genus-$g$ generalisation (\cref{kPlanarProduct,gkPlanarProduct}).  
  \item \Cref{examples} uses \cref{Structure} to derive product structure theorems for string graphs, powers of bounded-degree graphs, and $k$-nearest neighbour graphs.
  \item \Cref{FramedSection} proves a product structure theorem for $d$-framed multigraphs in which the graph $H$ has treewidth $3$ and is planar.  Using this result, we derive product structure theorems for $1$-planar graphs and map graphs, again with a planar graph $H$.  
  \item \Cref{Applications} quickly surveys applications of product structure and the consequences of the current work.  
  \item \cref{WrappingUp} concludes with an open problem.
\end{compactitem}

@@

The purpose of this section is to prove \cref{ShortcutPartition} and its specialization to $k$-planar graphs \cref{kPlanarProduct}.

\cref{ShortcutPartition} shows how, given a $(k,d)$-shortcut system $\SS$ of a graph $G$, an $H$-partition of $G$ can be used to obtain a $J$-partition of $G^{\SS}$ where the layered width  does not increase dramatically and the treewidth of $J$ is not much more than the treewidth of $H$.  For convenience, it will be helpful to assume that $\SS$ contains a length-1 $vw$-shortcut for every edge $vw\in E(G)$.  Since $G^\SS$ is defined to be a supergraph of $G$, this assumption has no effect on $G^{\SS}$ but eliminates special cases in some of our proofs.

@@

% We are now ready to complete the proof.
% , which we restate here for convenience:

@@

\subsection{\boldmath $k$-Planar Graphs}

@@

As mentioned in the introduction, product structure theorems have been established for several minor-closed classes in addition to planar graphs, including the following version of \cref{PlanarProduct} for graphs of bounded Euler genus.

\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]\label{GenusProduct}
  Every graph of Euler genus $g$ is contained in:
  \begin{compactenum}[(a)]
  \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
  \end{compactenum}
\end{thm}

The definition of $k$-planar graphs naturally generalises to genus-$g$ surfaces. A graph $G$ embedded on a surface $\Sigma$ is $(\Sigma,k)$-plane if every edge of $G$ is involved in at most $k$ crossings.  A graph $G$ is $(g,k)$-planar if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$. \cref{AddDummy} immediately generalises as follows:

Theorems~\ref{ShortcutProduct}  and \ref{GenusProduct}(b) imply a product structure theorem for $(g,k)$-planar graphs. The resulting bounds are improved by the following theorem, which is proved using exactly the same approach used in the proof of \cref{kPlanarProduct} (applying \cref{GenusProduct}(b) instead of \cref{PlanarProduct}(b)). We omit repeating the details.

\section{Graphs Derived from Shortcut Systems}
\label{examples}

In this section we give three examples of well-known graph families that can be expressed in terms of shortcut systems and which therefore have product structure theorems.

@@

A similar proof to that of \cref{StringShortcut} shows that every $(g,\delta)$-string graph is $(g,2\delta^2)$-planar. Theorems~\ref{PlanarProduct}(b), \ref{ShortcutProduct}  and \ref{GenusProduct}(b):
 % and \cref{StringShortcut,qn} imply:

@@

\section{Framed Graphs, $1$-Planar Graphs, and Map Graphs}
\label{FramedSection}

@@

For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it has an embedding in $\R^2$ having a biconnected plane spanning (multi)subgraph $G_0$, whose facial cycles each have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$. The graph $G_0$ is called the \defin{frame} of $G$. This definition (for simple graphs) was introduced by \citet{BDGGMR}.

Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar multigraph (by adding a vertex inside each face $F$ of $G_0$ adjacent to the vertices of $F$). Thus \cref{ShortcutProduct} is applicable. However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}

@@

\cref{d_framed_product_stucture} is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs.  For simple 1-planar graphs it is essential that we allow frames with parallel edges.

@@

\note{PM}{I would also like to change $G'$ to $G_0$ and $N'$ to $N_0$.  One issue is holding me back, though: We recurse on cycles $F_1,\ldots,F_m$ that define subgraphs $N_1,\ldots,N_m$ of $G$ and subgraphs $N_1',\ldots,N_m'$ of $G'$.  Will these latter subgraphs becomes $N_{0,1},\ldots,N_{0,m}$?}

% \note{DW}{We need to define ``closure''}

@@

% $1$-planar graphs are of particular interest to the graph drawing community~\citep{kobourov.liotta.ea:annotated}. In this case, we obtain better constants and an additional property (planarity) of $H$.

@@

\begin{lem}\label{1_planar_is_4_framed}
A multigraph $G$ is $1$-planar if and only if it is contained in a $4$-framed multigraph.
\end{lem}

% \note{DW}{Write  ``A multigraph $G$ is $1$-planar if and only if ...''}

@@

\begin{proof}
Let $G$ be a $1$-plane multigraph. We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.  This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.

@@

\note{PM}{Check David's commit: a0e20aff83d12c74127331854ce9fb75e002efe3}

@@

@@

As discussed in the introduction, product structure has been used to resolve a number of problems on planar graphs.  In most cases, these results hold for any graph with product structure. In this section, we survey some of the consequences of this for the graph classes considered in the previous two sections.

% Here we discuss some of the consequences of the above theorems for $k$-planar and $(g,k)$-planar graphs.
% 
% \referee{1}{4. Sections 4 and 5: I think as they are currently organized they distract from the purpose of the paper, as stated in the introduction, to ``prove product structure theorems for several non-minor-closed classes of interest.'' Which of Corollaries 1--4 and Theorems 12--17 are most important? I would remove any mention of the applications from the examples section, put that section first, and state in that section a single main theorem with all of the best bounds on the product structure. The applications section is mostly a survey and I think should be focused on the most exciting new corollaries. Theorems 18 and 19 should be in the applications section.}
% 
% \note{DW}{I don't want to do a complete restructuring. In our response, let's just say why we structure it the way we do.}
% \note{PM}{Sure, though I think they proposing a chop rather than a restructuring.}

@@

Since all the product structure theorems presented thus far upper bound the treewidth of $H$, 
\cref{qn} immediately implies bounds on the queue number of all graphs in these classes.  An improvement can be made in the cases where $H$ is planar and has treewidth $3$, as follows.  
\citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$. Thus the graph $H$ in \cref{1_planar_product} has queue-number at most $5$. \cref{qn,1_planar_product,NewMapGraph} then imply:

@@

\begin{cor}
\label{dMapQueue}
Every $d$-map graph has queue-number at most $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$.
\end{cor}

% \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
% \note{OM}{Ok}
% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
% \note{PM}{Ok}

The following corollary summarizes the results for other graph classes.
\begin{cor}\label{q_cor} 
  The following bounds hold for the queue number of any graph from each of the following classes:
  \begin{compactenum}
    \item $k$-planar: $2^{O(k^3)}$
    \item $(g,k)$-planar: $g2^{O(k^3)}$ 
    \item $(g,\delta)$-string: $g2^{O(\delta^3)}$
    \item $1$-planar: $115$
    \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$
    \item $(g,d)$-map: $g2^{O(d)}$
    \item $k$-NN: $2^{O(k^6)}$
    % \item $(g,d)$-map graph
    % \item 
  \end{compactenum}
\end{cor}

Note that \citet{DJMMUW20} previously proved the bound of
$O(g^{k+2})$ for $k$-planar graphs using \cref{GenusProduct} and an ad-hoc method. Our result provides a better bound when $g>2^{k^2}$.

@@

% \cref{non-repetitive,kPlanarProduct,1-planar,gkPlanarProduct} imply the following result:
% 
% \begin{cor}\quad
% \begin{compactitem}
% \item	For every $1$-planar graph $G$, $\pi(G)\le 7\times 4^4=1792$.
% %\item For every $k$-planar graph $G$,
% %	$\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
% \item	For every $(g,k)$-planar graph $G$,
% 	$  \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}.$
% % \note{DW}{I have changed $4^{\binom{k+5}{4}}$ to $4^{\binom{k+4}{3}}$, okay?}
% % \note{PM}{Ok}
% \end{compactitem}
% \end{cor}

The following corollary summarizes the results for other graph classes.
\begin{cor}\label{q_cor} 
  The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
  \begin{compactenum}
    \item $k$-planar: $(18k^2+48k+30) 4^{\binom{k+4}{3}}$
    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}$ 
    \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
    \item $1$-planar: $1792$
    \item $d$-map: $24 (d+3\floor{d/2}-3) \,  - \ceil{\tfrac{3}{2} (d+3\floor{d/2}-3)}$
    \item $(g,d)$-map: $g4^{O(d)}$
    \item $k$-NN: $4^{O(k^6)}$
    % \item $(g,d)$-map graph
    % \item 
  \end{compactenum}
\end{cor}

@@

\section{Orphaned Section on Structure}

Prior to this work, the strongest structural description of $k$-planar or $(g,k)$-planar graphs (or any of the other classes presented in \cref{examples}) was in terms of layered treewidth, which we now define.  A \defin{layered tree-decomposition} $(\mathcal{L},\mathcal{T})$ consists of a layering $\mathcal{L}$ and a tree-decomposition $\mathcal{T}$ of $G$. The layered width of $(\mathcal{L},\mathcal{T})$ is $\max\{|L\cap B|: L\in \mathcal{L},\, B\in \mathcal{T}\}$.  The \defin{layered treewidth} of $G$ is the minimum layered width of any layered tree-decomposition of $G$. \citet{dujmovic.morin.ea:layered} proved that planar graphs have layered treewidth at most 3, that graphs of Euler genus $g$ have layered treewidth at most $2g+3$, and more generally that a minor-closed class has bounded layered treewidth if and only if it excludes some apex graph. \citet{dujmovic.eppstein.ea:structure} show that every $k$-planar graph has layered treewidth at most $6(k+1)$, and more generally that every $(g,k)$-planar graph has layered treewidth at most $(4g+6)(k+1)$. It follows from this result that $(g,k)$-planar graphs have treewidth $O(\sqrt{(g+1)(k+1)n})$ and thus have balanced separators of the same order, which can also be concluded from the work of \citet{FP08}. In related work, \citet{grigoriev.bodlaender:algorithms} used structural results to obtain approximation algorithms for $(g,k)$-planar graphs, and \citet{PachToth97} determined the maximum number of edges in a $k$-planar graph (up to a constant factor).

%%%%%%%%%%%%%%%%%%%
\label{sec-k-planar}

@@

% As mentioned above, product structure theorems have been established for several minor-closed classes in addition to planar graphs. The first generalises \cref{PlanarProduct} for graphs of bounded Euler genus.
% 
% \begin{thm}[\citep{DJMMUW20,UWY,DHHW}]
% \label{GenusProduct}
% Every graph of Euler genus $g$ is contained in:
% \begin{compactenum}[(a)]
% \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
% \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
% \end{compactenum}
% \end{thm}
